```{r}

library(tidyverse)
library(janitor)
library(janeaustenr)
library(tidytext)
library(textdata)

```


```{r}

austen_books() %>% 
  count(book)

pp_ss <- austen_books() %>% 
  filter(book %in% c("Sense & Sensibility", "Pride & Prejudice")) %>% 
  unnest_tokens(word, text)

```
Below are the most common 5 words in _sense and sensibility_ and _pride and prejudice_. The order does change but in both books the 5 words are;
* to
* the
* of
* and
* her

```{r}

most_pop <- pp_ss %>% 
  group_by(book) %>% 
  count(word) %>% 
  slice_max(n, n = 5)

most_pop
```

The most common non-stop words in these 2 books are below;
This appers to be a series of names and the word time.

```{r}
pp_ss %>% 
  anti_join(stop_words) %>% 
  group_by(book) %>% 
  count(word, sort = TRUE) %>% 
  slice_max(n, n = 5)
```

The top sentiment words for each book are listed below. One for each book is miss which could be part of a name.

```{r}
pp_ss %>% 
  anti_join(stop_words) %>% 
  left_join(get_sentiments("afinn")) %>% 
  drop_na() %>% 
  group_by(book) %>% 
  count(word) %>% 
  slice_max(n, n = 5)
```

Across the two books there are around 450 occurrences where Miss is capitalized and appears to be used as part of a name. There are only around 500 instances of the word miss in these books so this would not be one of the top 5 results without it's other purpose as part of a name. Miss will therefore be removed from the previous sentiment analysis to get a more accurate answer. 

```{r}

austen_books() %>% 
  filter(book %in% c("Sense & Sensibility", "Pride & Prejudice")) %>% 
  unnest_tokens(bigram, text, token = "ngrams", n = 2, to_lower = FALSE) %>% 
  mutate(miss = str_detect(bigram, "^Miss")) %>% 
  filter(miss == TRUE) %>% 
  count(book)
  

```

Now there is a more accurate table of sentiments portrayed in these two books.

```{r}

pp_ss %>% 
  anti_join(stop_words) %>% 
  left_join(get_sentiments("afinn")) %>% 
  drop_na() %>% 
  filter(word != "miss") %>% 
  group_by(book) %>% 
  count(word) %>% 
  slice_max(n, n = 5)

```