---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```
__Made assumption that all weights are in kilograms__

# Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}

backpack <- CodeClanData::backpack
backpack
```
There a 100 variables in this data set. 
These comprise backpack_weight, body_weight, back problems, year and units as integers.
Ratio is backpack_weight/body_weight as a numeric value.
Major, sex and status are factors.
Unsure what units refers to at all and uncertain what different status will relate to.

Back problems looks like 0 and 1 values and may be better as logical.

# Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}

backpack %>% 
  ggplot()+
  geom_bar(aes(x = backpack_weight))+
  labs(title = "Number of students with a set backpack weight",
       y = "Number of students",
       x = "Backpack weight")+
  scale_x_continuous(breaks = 1:38)+
  scale_y_continuous(breaks = c(0,2,4,6,8,10,12))

backpack %>% 
  ggplot()+
  geom_boxplot(aes(x = backpack_weight))+
  labs(title = "Boxplots of students with a set backpack weight",
       x = "Backpack weight")+
  scale_x_continuous(breaks = 1:38)
```

This is one variable (backpack_weight) by a count of iteself therefore a bar/col graph would be useful. Histogram may also have been an option.
There are 3 students with a backpack weight of only 2kg.
There is 1 students with a backpack weight of 35kg.

By plotting a box plot in addition it is clear that 50% of students have a backpack weighing between 8kg and 14kg with some statistical outliers on the upper end of the data (>22kg).The average weight is 11kg.



# Question 3.

Are there differences in backpack weight between sex?

```{r}
backpack %>% 
  ggplot()+
  geom_bar(aes(x = backpack_weight))+
  facet_grid(~sex)+
    labs(title = "Number of students with a set backpack weight by gender",
       y = "Number of students",
       x = "Backpack weight")+
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,25))+
  scale_y_continuous(breaks = c(0,2,4,6,8,10,12))

backpack %>% 
  ggplot()+
  geom_boxplot(aes(backpack_weight))+
    facet_grid(~sex)+
  labs(title = "Boxplots of students with a set backpack weight",
       x = "Backpack weight")+
  scale_x_continuous(breaks = 1:38)
```



Question 4.

What is the relationship between body weight and backpack weight?

Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

Question 6.

What proportion of students report back pain?

Question 7.

Is this proportion different between men and women?

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

3 Extensions


Question 1.

Does reported backpain differ across the year groups? Ignore year 0 in your data.

Question 2.

Is there a difference between how heavy the bags those in different years carry?

