---
title: "R Notebook"
output: html_notebook
---

library(tidyverse)
library(CodeClanData)

Question 1.

Load in the dataset and have a look at it. What are the dimensions, variable types, variable names, etc.?

_Dimensions 18 columns (variables), 242 rows (observations). Types - chr, dbl. Names - see list below._

```{r}
drinks_content
```


Question 2.

Let’s first investigate the calories of different drinks. Select the variables Beverage_category, Beverage, Beverage prep and Calories from your data. Since we are interested in the calorie content, check if there are any NA values in the data, and drop them if there are.

_Selected relevant columns_
_Checked for n/a values within the Calories row._

```{r}
drinks <- drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  mutate(Calories = Calories, na.rm = TRUE)
drinks
```





Question 3.

Create a new variable (column) called calorie_diff, which stores the difference between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. (hint: you’ll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

_Table below_

```{r}
drinks %>% 
  mutate(calories_diff = Calories - 135)
drinks
```




Question 4.

Summarise the mean number of calories in each beverage_category. Which 3 category of drinks have the most calories? Which 3 drinks have the least? Write a small summary of your findings.

_Grouped by category, summarise average value per category, then 2 functions to arrange asc and desc._
_Lowest - Coffee, Shaken Iced Beverages, Classic Espresso Drinks_
_Highest - Smoothies,Frappuccino® Blended Coffee, Signature Espresso Drinks_


```{r}
drinks %>% 
  group_by(Beverage_category) %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
  arrange(average_value)
```
```{r}
drinks %>% 
  group_by(Beverage_category) %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(average_value))
```


Question 5.

Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?

_Same thing but different column to group on._

```{r}
drinks %>% 
  group_by(Beverage_prep) %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(average_value))
```



Question 6.

Which Beverage_prep type contains more than the average calories of all drinks?
Hint

_All drinks first - average value is 193.8719_
```{r}
drinks %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(average_value))
```
_Those listed below have a higher calorie value than average._

```{r}
drinks %>% 
  group_by(Beverage_prep) %>% 
  summarise(average_value = mean(Calories, na.rm = TRUE)) %>% 
  arrange(desc(average_value)) %>% 
  filter(average_value >193.8719)
```


Question 7.

Which is the best type of coffee drink (and by ‘drink’ we mean the combination of Beverage_category, Beverage and Beverage_prep) to get if you’re worried about consuming too many calories?

_Tazo Tea, has no calories if this counts_
_Best coffee is brewed Coffee, size doesn't matter much_

```{r}
drinks2 <- drinks_content %>% 
  mutate(Calories = Calories, na.rm = TRUE) %>% 
  arrange(Calories)
drinks2
```


