---
title: "R Notebook"
output: html_notebook
---
load libraries
```{r}
library(tidyverse)
library(janitor)
```
load data
```{r}
revenue_breakdown <- clean_names(CodeClanData::qb_revenue_breakdown)
monthly_sales <- clean_names(CodeClanData::qb_monthly_sales)
competitors <- clean_names(CodeClanData::qb_competitors)
device_data <- clean_names(CodeClanData::qb_device_data)
```
create theme
__Couldn't work out how to make lines faint but used a fainter colour instead__
```{r}
my_theme <- theme(panel.background = element_rect(fill = "white"),
                  axis.line = element_line(size = 1),
                  panel.grid.major = element_line(colour = "azure2", size = 0.2),
                  text = element_text(size = 12),
                  title = element_text(size = 16, hjust = 0.5),
                  axis.title = element_text(size = 8))

col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```

# Question 1.
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.
__First the question I think is actually being asked__
__Then the question I think is meant__

```{r}

revenue_breakdown %>% 
  ggplot(aes(x = year, fill = product))+
  geom_bar()+
  labs(title = "Different revenue sources by year")+
  scale_fill_manual(values = col_scheme)+
  my_theme


revenue_breakdown %>% 
  ggplot(aes(x = year, y = revenue, fill = product))+
  geom_col()+
  labs(title = "Revenue generated by different sources across years")+
  scale_fill_manual(values = col_scheme)+
  my_theme

```


Question 2.
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}

monthly_sales %>% 
  filter(cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot(aes(x = date, y = amount))+
  geom_line()+
  labs(title = "Monthly sales of Kwikbit Unit 1")+
  scale_fill_manual(values = col_scheme)+
  my_theme


```

Question 3.
Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.


```{r}

monthly_sales %>% 
  ggplot()+
  geom_line(aes(x = date, y = amount, colour = cashflow))+
  labs(title = "Monthly sales and revenue breakdown")+
  scale_fill_manual(values = col_scheme)+
  my_theme

```

Question 4.
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.


```{r}

competitors %>% 
  ggplot(aes(x = year, y = revenue, fill = company))+
  geom_area()+
  labs(title = "Annual sales of fitness trackers by competitors")+
  scale_y_continuous(labels = scales::comma)+
  scale_fill_manual(values = col_scheme)+
  my_theme


competitors %>% 
  ggplot(aes(x = year, y = revenue, fill = company))+
  geom_area()+
  labs(title = "Annual sales of fitness trackers by competitors")+
  scale_y_continuous(labels = scales::comma)+
  facet_wrap(~company, ncol = 2)+
  scale_fill_manual(values = col_scheme)+
  my_theme

```


Question 5.
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.


```{r}

competitors %>% 
  ggplot(aes(x = year, y = revenue, group = company))+
  geom_line()+
  geom_point()+
  labs(title = "Annual revenue of competitors")+
  scale_fill_manual(values = col_scheme)+
  my_theme

```



Question 6.
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
device_data

device_data %>% 
  mutate(time = hours+(mins/60)) %>% 
  ggplot(aes(x = time, y = counts))+
  geom_line()+
  labs(title = "Steps tracked by device and person")+
  facet_wrap(~id + device, ncol = 4)+
  scale_fill_manual(values = col_scheme)+
  my_theme

```


2.2 Part 2
Question 1.
Take the plots that you produced in part one and now polish them by:

Adding appropriate labels
Changing the scales and coordinates when appropriate.
Applying a unified theme, which is described below:
Graphs should have white backgrounds, and use colour sparingly.
There should be faint grid lines.
Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller.
All plots should use colours from the following company colour scheme.


col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")


3 Extension
Question 1.
Try out different ways of plotting these three datasets. Which graphs do you think are particularly effective at showing QikBit’s success?

