
```{r}
library(tidyverse)
library(factoextra)
library(janitor)
library(ggrepel)
```
# Clustering homework
# You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

Customer id is just a sequential value.
Gender is split 112 Female, 88 Male.
Age ranges from 18 to 70, median of 36 and a right skew to the data.

Annual income ranges from 15k to 137k, median of 62k and again a right skew to the data.

Spending score ranges from 1 to 99 with an median score of 50. Data has multiple peaks visible in histogram but appears quite normal overall through boxplot. 
```{r}
customers <- read_csv("data/mall_customers.csv") %>% 
  clean_names() %>% 
  select(-customer_id)

summary(customers)

customers %>% 
  count(gender)

customers %>% 
  ggplot(aes(x = age))+
  geom_boxplot()

customers %>% 
  ggplot(aes(x = annual_income_k))+
  geom_boxplot()

customers %>% 
  ggplot(aes(x = spending_score_1_100))+
  geom_histogram()+
  geom_boxplot()
```
When plotting the data there appear to be 5 clusters. Middle income and spending. Then 4 with low/high income against low/high spending.
```{r}

customers %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100))+
  geom_point()

```
# We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

Using the 3 graphs below it appears as though there may be 5 clusters though the line for gap analysis does not appear to be in the expected location.

```{r}

customers_scaled <- customers %>% 
  select(annual_income_k, spending_score_1_100) %>% 
  mutate(across(everything(), scale))

fviz_nbclust(customers_scaled,
             kmeans,
             method = "wss",
             nstart = 25)

fviz_nbclust(customers_scaled,
             kmeans,
             method = "silhouette",
             nstart = 25)

fviz_nbclust(customers_scaled,
             kmeans,
             method = "gap",
             nstart = 25)
```


# Perform k-means clustering and chose a value of k.
Have chosen 5 based on the graphs above.
```{r}

clusters <- kmeans(customers_scaled,
                   centers = 5,
                   nstart = 25)
clusters
```


# Visualise the clustering for your chosen value of k.


```{r}
customers_clustered <- customers %>% 
  mutate(cluster = as.factor(clusters$cluster))

customers_clustered
```

```{r}

customers_clustered %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = cluster), colour = c(2:6))+
  geom_point()
```


# Do you think the clustering seems a good fit for this data?

Yes - based on the visualization above the data appears to fit very nicely for clustering.

# Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

Comments via labels in the plots below. Should look to attract more business in high_income, high_spend category and see how to get high_income, low_spend to spend more.

```{r}
customers_labeled <- customers_clustered %>% 
  mutate(label = case_when(
    cluster == 1 ~ "Mid income, mid spend",
    cluster == 2 ~ "Low income, low spend",
    cluster == 3 ~ "Low income, high spend",
    cluster == 4 ~ "High income, low spend",
    TRUE         ~ "High income, high spend"
  ))

customers_labeled
```

```{r}

label_df <- customers_labeled(cluster = levels(customers_labeled$label),
                              label = paste0(levels(customers_labeled$label)))

label <- customers_labeled %>% 
  group_by(label) %>% 
  summarise(x = min(annual_income_k), y = max(spending_score_1_100)) %>% 
  left_join(label_df)

customers_labeled %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = label), colour = c(2:6))+
  geom_point(show.legend = FALSE)+
  geom_text(aes(x = labs$meanai, y = labs$meanss, label = labs$label),
            show.legend = FALSE,
            check_overlap = TRUE,
            size = 4,
            colour = "black")
```

