```{r}
library(tidyverse)
library(gtools)
library(janitor)
source("prob.R")
```

# Question
Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

```{r}
titanic <- read_csv("data/titanic.csv")
titanic %>% 
  count(class)

new_titanic <- titanic%>% 
  mutate(pass_crew = if_else(class %in% c("1st", "2nd", "3rd"), "passenger", "crew"))

new_titanic %>% 
  count(pass_crew, survived)
```
```{r}
#probability for crew to survive - 0.237 (23.7%)
crew_titanic <- new_titanic %>% 
  filter(pass_crew == "crew")

crew_survival <- crew_titanic %>% 
  filter(survived == "yes") %>% 
  summarise(prob = n()/nrow(crew_titanic))
crew_survival


#probability for passenger to survive - 0.379 (37.9%)
passenger_titanic <- new_titanic %>% 
  filter(pass_crew == "passenger")

passenger_survival <- passenger_titanic %>% 
  filter(survived == "yes") %>% 
  summarise(prob = n()/nrow(passenger_titanic))
passenger_survival

#Same results but method to get data into table if further work required.
new_titanic %>% 
  group_by(pass_crew) %>% 
  mutate(number_in_group = n()) %>% 
  ungroup() %>% 
  group_by(pass_crew, survived) %>% 
  mutate(survival_status_in_group = n()) %>% 
  mutate(survival_status_chance = survival_status_in_group/number_in_group) %>% 
  distinct(survival_status_chance) %>% 
  filter(survived == "yes")

#
new_titanic %>% 
  group_by(class) %>% 
  mutate(number_in_class = n()) %>% 
  ungroup() %>% 
  group_by(class, survived) %>% 
  mutate(survival_status_in_class = n()) %>% 
  mutate(survival_status_chance_class = survival_status_in_class/number_in_class) %>% 
  distinct(survival_status_chance_class) %>% 
  filter(survived == "yes") %>% 
  arrange(desc(survival_status_chance_class))
  
```


# Question 1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?
_Without code - 4/6_
_With code - The answer is 0.666667 or 4/6_
```{r}
one_dice <- rolldie(1)

chance_of_under_five <- one_dice %>% 
  filter(X1 <5) %>% 
  summarise(probability = n()/nrow(one_dice))
chance_of_under_five
```

# Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?
_Without code - 6^3 = 216_
_With code - there are 216 possible outcomes in the table_
```{r}
rolldie(3)
```


# Question 3.
In an experiment where we roll four dice:

how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)?
what is the probability of this event?
what assumptions do we apply in answering part ii?
Hint
For part i, rowSums(.) may help you setup up the necessary random variable total (or you can use pivot_longer())
_Without code - There are now 6^4 outcomes = 1296, there are 18 outcomes, 18/1296 probability _
_With code there are 1296 outcomes and 34 of these produce a total of 5,6 or 7. There is a 0.026 (2.6%) probability of these results_
_Assumption is made that the total is all that is important and the dice are all equivalent. Also all dice a fair._
```{r}

four_die <- rolldie(4)

four_die %>% 
  mutate(sum_of_rolls = rowSums(.)) %>% 
  filter(sum_of_rolls %in% c(5,6,7)) %>% 
  summarise(probability = n()/nrow(four_die))
```

# Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint
This is an arrangement where order matters. Of all the possible arrangements, how many are correct?
_Without code 5*4*3*2, only one outcome in order so 1/120_
_With code, 120 outcomes, still only 1 correct answer 0.00833 (0.8%)_


```{r}

people <- c(1,2,3,4,5)

permutations(5,5,people) %>% 
  as_tibble() %>% 
  summarise(probability = 1/n())
```



# Question 5.
I roll six dice:

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
Hint
Have a look at the isin() function in the prob.R file. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE)
_Without code, there is one possibility for in order out of 6^6 = 46656 possible outcomes. Would think there are 6*5*4*3*2 = 720 outcomes with all numbers shown_
_With code, calculates to 46656 outcomes. This also calcualtes 720 outcomes where this is true. 720/46656 = 0.154 (1.5%)_


```{r}

six_die <- rolldie(6)

isin(six_die, c(1,2,3, 4, 5, 6), ordered = FALSE) %>%
  as.tibble() %>% 
  count(value)

```


# Question 6.
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

Hints
Investigate the cards() function in the prob.R file, if you’d like to try this in R (make a deck without jokers).
You could think about adding your own black_card random variable to the sample space, which would be TRUE if the suit is ‘Spade’ or ‘Club’, and FALSE otherwise.
_Without code 0.5 (50%)_
_With code - can't see the way to work this one_
```{r}





cards <- function (jokers = FALSE)
{
  x <- c(2:10, "J", "Q", "K", "A")
  y <- c("Club", "Diamond", "Heart", "Spade")
  z <- c("Black", "Red")
  res <- expand.grid(rank = x, suit = y)
  if (jokers) {
    levels(res$rank) <- c(levels(res$rank), "Joker")
    res <- rbind(res, data.frame(rank = c("Joker", "Joker"),
                                 suit = c(NA, NA)))
  }
  return(res)
}




```
