---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(ggfortify)
library(GGally)
```
# Data familiarisation
```{r}
avocados <- read_csv(here::here("data/avocado.csv")) %>% 
  clean_names() %>% 
  rename(small_sold = x4046, large_sold = x4225, x_large_sold = x4770)
# data is over 4 years
avocados %>% 
  count(year)
# no NA data in region but 54 variables, some appear to be region summaries rather than more localised
avocados %>% 
  count(region)

summary(avocados)
```

# Data cleaning for analysis

Some of these created columns will be alias for each other would only want to model on 1 ratio column or 1 sold column plus total fine on top.
```{r}
# create ratio for three sizes of avocado sales and bags

avocados_tidy <- avocados %>% 
  mutate(month = month(date, label = TRUE, abbr = FALSE)) %>% 
  mutate(ratio_small_sold = small_sold/total_volume,
         ratio_large_sold = large_sold/total_volume,
         ratio_x_large_sold = x_large_sold/total_volume,
         ratio_small_bags = small_bags/total_bags,
         ratio_large_bags = large_bags/total_bags,
         ratio_x_large_bags = x_large_bags/total_bags) %>% 
  select(-c(date, x1))

summary(avocados_tidy)

avocados_tidy_no_region <- avocados_tidy %>% 
  select(-region)

write_csv(avocados_tidy_no_region, here::here("data/clean_avocado.csv"))
```


# Initial modelling

_Small_sold, small_bags and total volume to explore_

```{r, message = FALSE}
avocados_tidy %>% 
  select(average_price, total_volume:x_large_bags) %>% 
  ggpairs()
```
_Type is definately worth looking at, year and month might be of interest_

```{r, message = FALSE}
avocados_tidy %>% 
  select(average_price, type, year, month) %>% 
  ggpairs()
```
_ratio of small sold and ratio of xl bags may be of interest_

```{r, message = FALSE}
avocados_tidy %>% 
  select(average_price, ratio_small_sold:ratio_x_large_bags) %>% 
  ggpairs()
```

Chosen variable does not appear significantly more normally distributed when log applied (right skewed data). Will retain as normal value for simplicity.

```{r}
avocados_tidy %>% 
  ggplot(aes(x = log(average_price)))+
  geom_histogram()

avocados_tidy %>% 
  ggplot(aes(x = average_price))+
  geom_histogram()
```

