---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(ggfortify)
library(GGally)
library(leaps)
library(modelr)
```

```{r}
avocados_tidy_no_region <- read_csv(here::here("data/clean_avocado.csv"))
```


```{r}
automodel <- regsubsets(average_price ~.,
                                 data = avocados_tidy_no_region,
                                 nvmax = 8,
                                 method = "exhaustive")

sum_forward <- summary(automodel)

sum_forward
```

As an exploratory glance models including type, ratio_large_bags, ratio_large_sold, ratio_small_sold, month and year.

```{r}
automodel_region <- regsubsets(average_price ~.,
                                 data = avocados_tidy,
                                 nvmax = 8,
                                 method = "forward")

sum_forward2 <- summary(automodel_region)

sum_forward2
```


When leaving region in only Houston was identified as an input by the automodel.

# 1st round of modeling

_Type has a 0.38 r^2. A good starting point would be to build a model based on type_
```{r}

model1a <- lm(average_price ~ type, avocados_tidy_no_region)
summary(model1a)

model1b <- lm(average_price ~ ratio_small_sold, avocados_tidy_no_region)
summary(model1b)

model1c <- lm(average_price ~ ratio_large_sold, avocados_tidy_no_region)
summary(model1c)

model1d <- lm(average_price ~ total_volume, avocados_tidy_no_region)
summary(model1d)

model1e <- lm(average_price ~ year, avocados_tidy_no_region)
summary(model1e)
```
Autoplot doesn't help much for a 2 option variable but normal Q-Q looks normal enough.

```{r}

autoplot(model1a)

```
From this part of the ggpairs on the residual data it would be good to look at small sold, large bags and total_volume

```{r, message = FALSE}
price_resid <- avocados_tidy_no_region %>% 
  add_residuals(model1a) %>% 
  add_predictions(model1a) %>% 
  select(-average_price, -type)

price_resid %>% 
  select(resid, total_volume:x_large_bags) %>%
  ggpairs()

```

4 very promising categories here for ratio small/large and bags small/large.

```{r, message = FALSE}

price_resid %>% 
  select(resid, year:ratio_large_bags) %>% 
  ggpairs()

```

# 2nd round of modeling

```{r}

model2a <- lm(average_price ~ type + ratio_large_sold, avocados_tidy_no_region)
summary(model2a)

model2a <- lm(average_price ~ type + ratio_small_sold, avocados_tidy_no_region)
summary(model2a)

model2a <- lm(average_price ~ type + ratio_small_bags, avocados_tidy_no_region)
summary(model2a)

model2a <- lm(average_price ~ type + ratio_large_bags, avocados_tidy_no_region)
summary(model2a)

```

```{r}

autoplot(model2a)

```


Best additional item to fit to the data is ratio_large_sold. This can be an alias for ratio small_sold and ratio_x_large_sold, if one of these is added as a best fit the other must be removed. Likewise large_sold may not add much with this variable already in place.

```{r, message = FALSE}
price_resid_2 <- avocados_tidy_no_region %>% 
  add_residuals(model2a) %>% 
  add_predictions(model2a) %>% 
  select(-average_price, -type, - ratio_large_sold)

price_resid_2 %>% 
  select(resid, total_volume:x_large_bags) %>%
  ggpairs()

```


```{r, message = FALSE}

price_resid_2 %>% 
  select(resid, year:ratio_x_large_bags) %>% 
  ggpairs()

```

# 3rd round of modeling

```{r}

model3a <- lm(average_price ~ type + ratio_large_sold + ratio_small_sold, avocados_tidy_no_region)
summary(model3a)

model3b <- lm(average_price ~ type + ratio_large_sold + year, avocados_tidy_no_region)
summary(model3b)

model3c <- lm(average_price ~ type + ratio_large_sold + month, avocados_tidy_no_region)
summary(model3c)

model3d <- lm(average_price ~ type + ratio_large_sold + small_sold, avocados_tidy_no_region)
summary(model3d)

model3e <- lm(average_price ~ type + ratio_large_sold + x_large_sold, avocados_tidy_no_region)
summary(model3e)

```

Model3c adding month looks to be the next best predictor

```{r}

autoplot(model3c)

price_resid_2 %>% 
  ggplot(aes(x = month, y = resid))+
  geom_point()

coplot(resid ~ month | year,
       panel = function(x,y, ...){
         points(x,y)
         abline(lm(y ~x), col = "blue")
       },
       data = price_resid_2, rows = 1)

model3c2 <- lm(average_price ~ type + ratio_large_sold + month:year, avocados_tidy_no_region)
summary(model3c2)
#maybe month:year is better than month alone, will test this in next set

```

```{r, message= FALSE}
price_resid_3 <- avocados_tidy_no_region %>% 
  add_residuals(model3c) %>% 
  add_predictions(model3c) %>% 
  select(-average_price, -type, - ratio_large_sold, -month)

price_resid_3 %>% 
  select(resid, total_volume:large_bags) %>%
  ggpairs()
```


```{r, message= FALSE}
price_resid_3 %>% 
  select(resid, x_large_bags:ratio_x_large_bags) %>% 
  ggpairs()
```

# 4th round of modeling

```{r}

model4a <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags, avocados_tidy_no_region)
summary(model4a)

model4b <- lm(average_price ~ type + ratio_large_sold + month + ratio_small_bags, avocados_tidy_no_region)
summary(model4b)

model4c <- lm(average_price ~ type + ratio_large_sold + month + year, avocados_tidy_no_region)
summary(model4c)

model4d <- lm(average_price ~ type + ratio_large_sold + month + small_sold, avocados_tidy_no_region)
summary(model4d)


```
Adding ratio_large_bags proved most effective which helps for simlicity as the amount of large_sold is already in the model

```{r, message= FALSE}
price_resid_4 <- avocados_tidy_no_region %>% 
  add_residuals(model4a) %>% 
  add_predictions(model4a) %>% 
  select(-average_price, -type, - ratio_large_sold, -month, -ratio_large_sold)

price_resid_4 %>% 
  select(resid, total_volume:large_bags) %>%
  ggpairs()
```


```{r, message= FALSE}
price_resid_4 %>% 
  select(resid, x_large_bags:ratio_x_large_bags) %>% 
  ggpairs()
```

# 5th round of modeling

```{r}

model5a <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year, avocados_tidy_no_region)
summary(model5a)

model5b <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + large_sold, avocados_tidy_no_region)
summary(model5b)

model5c <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + x_large_sold, avocados_tidy_no_region)
summary(model5c)

model5d <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + small_sold, avocados_tidy_no_region)
summary(model5d)

model5e <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + total_volume, avocados_tidy_no_region)
summary(model5e)

```
year is the best addition

```{r, message= FALSE}
price_resid_5 <- avocados_tidy_no_region %>% 
  add_residuals(model5a) %>% 
  add_predictions(model5a) %>% 
  select(-average_price, -type, - ratio_large_sold, -month, -ratio_large_sold, -year)

price_resid_5 %>% 
  select(resid, total_volume:large_bags) %>%
  ggpairs()
```


```{r, message= FALSE}
price_resid_5 %>% 
  select(resid, x_large_bags:ratio_x_large_bags) %>% 
  ggpairs()
```



# 6th round of modeling


```{r}
model6a <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + total_volume, avocados_tidy_no_region)
summary(model6a)

model6b <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + small_sold, avocados_tidy_no_region)
summary(model6b)

model6c <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold, avocados_tidy_no_region)
summary(model6c)

model6d <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + x_large_sold, avocados_tidy_no_region)
summary(model6d)

model6e <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + small_bags, avocados_tidy_no_region)
summary(model6e)

```

For simplicity rather than taking the best variable to add (small_sold) will continue to add large advocado variable as there is a minimal difference in this addition 0.5628 Vs 0.5626. Would rather retain some simplicity in the model.

```{r}

model6c <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold, avocados_tidy_no_region)
summary(model6c)

```

```{r, message= FALSE}
price_resid_6 <- avocados_tidy_no_region %>% 
  add_residuals(model6c) %>% 
  add_predictions(model6c) %>% 
  select(-average_price, -type, - ratio_large_sold, -month, -ratio_large_sold, -year, -large_sold)

price_resid_6 %>% 
  select(resid, total_volume:large_bags) %>%
  ggpairs()
```

```{r, message= FALSE}
price_resid_6 %>% 
  select(resid, x_large_bags:ratio_x_large_bags) %>% 
  ggpairs()
```


# 7th round of modeling

```{r}

model7a <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold + ratio_small_sold, avocados_tidy_no_region)
summary(model7a)

model7b <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold + ratio_x_large_bags, avocados_tidy_no_region)
summary(model7b)

model7c <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold + ratio_x_large_sold, avocados_tidy_no_region)
summary(model7c)

model7d <- lm(average_price ~ type + ratio_large_sold + month + ratio_large_bags + year + large_sold + x_large_bags, avocados_tidy_no_region)
summary(model7d)

```

Best addition was to add ratio_x_large_sold but this only added 0.0008 to the multiple r squared value. Call this the point where we have seem diminishing returns and focus on model up to this point. Model6c to be tested.


```{r}
broom::glance(model1a)
broom::glance(model2a)
broom::glance(model3c)
broom::glance(model4a)
broom::glance(model5a)
broom::glance(model6c)

```

adjusted r^2 increases throughout modeling process. AIC and BIC both decrease throughout modeling process indicate model is not overfit. 

```{r}
anova(model6c, model5a)
```

Annova shows that final model is significantly statistically improved for predicting average_price than the previous model.