---
title: "R Notebook"
output: html_notebook
---

#loading libraries and data

```{r}
library(tidyverse)
books <- read_csv("data/books.csv")
```

#General overview of data

_The data is 11123 rows by 13 columns._
_The column names are;_
[1] "rowid"              "bookID"            
 [3] "title"              "authors"           
 [5] "average_rating"     "isbn"              
 [7] "isbn13"             "language_code"     
 [9] "num_pages"          "ratings_count"     
[11] "text_reviews_count" "publication_date"  
[13] "publisher" 

_There are no na values within this dataset._

```{r}
dim(books)
names(books)
sum(is.na(books))
```
#Overview of dataset and column types

```{r}
head(books)
```

#Select data the has value to my analysis using select function - removed first 2 columns and ISBN data

```{r}
selected_books <- books %>% 
  select(title, authors, average_rating, language_code, num_pages, ratings_count, text_reviews_count, publication_date, publisher)
dim(selected_books)
```

#find most highest rating and most rated books

```{r}
selected_books %>% 
  arrange(desc(average_rating))

selected_books %>% 
  mutate(total_ratings = ratings_count + text_reviews_count) %>% 
  arrange(total_ratings)
```

#While there are no na values, there are 0 values. Should filter these out prior to checking any averages etc.

```{r}
#just code to pull later for specific views - to delete
no_zero_books <- books %>% 
  filter(average_rating != 0, num_pages != 0, ratings_count != 0, text_reviews_count != 0)
no_zero_books
```
#Selected distinct data by book title, this removed 775 entires from the data_set (11123 - 10348)

```{r}
selected_books %>% 
  distinct(title)

11123 - 10348
```
#Find the authors with the most books
```{r}
selected_books %>% 
  group_by(authors) %>% 
  summarise(books_by_author = n()) %>% 
  arrange(desc(books_by_author))

selected_books %>% 
  group_by(authors) %>%
  distinct(title) %>% 
  summarise(books_by_author = n()) %>% 
  arrange(desc(books_by_author))



selected_books %>% 
  filter(authors %in% c("Robert Jordan", "Michael Connelly"))

selected_books %>% 
  filter(authors == "Stephen King")

selected_books %>% 
  filter(authors == "P.G. Wodehouse")
```

