```{r}
library(tidyverse)
library(infer)
library(janitor)
```
```{r}
transactions <- read_csv("data/online_retail_subset.csv")

transactions
```
For the first section we are interested in the purchase of two particular items:

item A - ‘HEART OF WICKER SMALL’ (StockCode 22469)
item B - ‘LARGE CAKE TOWEL PINK SPOTS’ (StockCode 21110)

# Question 1.
Calculate the support for item A (this will be the support for a single item)

Item A was purchased 110 times, total in database is 30000.
Support for item A is 0.00366667
```{r}
transactions %>% 
  filter(StockCode == 22469)

support_A <- transactions %>% 
  filter(StockCode == 22469) %>% 
  summarise(prop = n()/nrow(transactions))
support_A
```
support for item B
only 14 rows of item B. Support for this item is 0.000466667
```{r}
transactions %>% 
  filter(StockCode == 21110)

support_B <- transactions %>% 
  filter(StockCode == 21110) %>% 
  summarise(prop = n()/nrow(transactions))
support_B
```
# Question 2.
Calculate the support and confidence for rule (A→B).

Support A->B = 0.00023333
Confidence A-> = 1.127273
```{r}
sup_A_B <- transactions %>% 
  filter(StockCode == 22469 | StockCode == 21110) %>% 
  count(InvoiceNo) %>% 
  filter(n == 2) %>% 
  summarise(bought_together = n()/nrow(transactions))
sup_A_B


conf_A_B <- sup_A_B/support_A
conf_A_B
```

# Question 3.
Calculate the lift for (A→B)
lift is 136
```{r}
lift_A_B <- sup_A_B / (support_A * support_B)

lift_A_B
```

```{r}
library(arules)
library(arulesViz)
```


```{r}
transactions_reformat <- transactions %>%
  select(InvoiceNo, Description) %>%
  na.omit()

write_csv(transactions_reformat, "transactions_reformat.csv")

apriori_format <- read.transactions("transactions_reformat.csv", format = "single", sep = ",", header = TRUE, cols = c("InvoiceNo", "Description"))

```

```{r}
inspect(head(apriori_format))
```

```{r}
itemFrequencyPlot(transactions, topN = 20, type = "absolute")
```

