```{r}
library(tidyverse)
library(here)
library(janitor)
```


```{r}
a <- read_csv(here("Data/opendata_inc1620comb_hb.csv"))
b <- read_csv(here("Data/opendata_inc1620comb_region.csv"))
c <- read_csv(here("data/opendata_inc1620comb_scotland.csv"))
d <- read_csv(here("data/opendata_inc9620_hb.csv"))
e <- read_csv(here("Data/opendata_inc9620_region.csv"))
f <- read_csv(here("Data/opendata_inc9620_scotland.csv"))
```
```{r}
a

a %>% 
  count(Year)
```
A - healthboard, cancer sites, sex, all data 2016 - 2020, instances per age group, incident rate per 100,000 per age group. EASR european average for age group, WASR world average for age group

```{r}
b

b %>% 
  count(Region)
```
Does not specify to NHS borders region

```{r}
c

c %>% 
  count(Country)
```
All just general Scotalnd data

```{r}
d
```
D - healthboard, cancer sites, sex, by year, total instances for all ages, EASR european average for age group, WASR world average for age group

```{r}
e
```
B crossed with D

```{r}
f
```
c corssed with D
